The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-03-18 src/ms.tex
Lim edit: Beginning - end of Sec 2.2

Within \astropypkg, development may be roughly split into several
categories: new features, intra- and inter-package interoperability,
improvements to numerical precision, accuracy, and reproducibility, performance
and documentation enhancements, and bug fixes. We discuss some of these in the subsections below.

@@

\astropy now supports reading and writing tables in the American Astronomical
Society Journals' Machine-Readable Table (\href{https://docs.astropy.org/en/stable/io/ascii/write.html#cds-mrt-format}{MRT}) format. This ASCII format has
long been missing and was added to \astropyTable\ I/O in a Google Summer of Code 2021 project.
In addition, \astropyTable\ may also read from and write
to \href{https://asdf-standard.readthedocs.io/en/latest/}{ASDF},\footnote{Starting in v5.1,
direct \astropypkg support for ASDF has been deprecated in favor of moving it
to the new \package{asdf-astropy} \citep{asdf-astropy} package to accommodate
the differing release schedules between ASDF and \astropypkg.}
\href{https://docs.astropy.org/en/stable/io/unified.html#table-io-parquet}{Parquet},
and \href{https://wwwastro.msfc.nasa.gov/qdp/}{QDP} formats.

@@

\subsection{Numerical Precision, Accuracy, \& Reproducibility}
\label{sec:core-features-precision_accuracy_repreducibility}

@@

% \begin{itemize}
% \item {\bf User forums:} The Astropy Project has historically maintained a number of avenues for users to seek help or gain access to the developer community. This includes the Astropy mailing list, the Python in Astronomy Facebook group, and the Astropy Slack workspace. The Moore Foundation grant allows this Slack space to
% be on a paid plan; additionally NumFOCUS has negotiated a special rate for
% open-source projects. While the Slack and Astropy-dev mailing lists are primarily used to discuss the project direction and updates, it was noted that the use of a Facebook community could present a barrier to open source. The Astropy Project identified a need for a public, archived, searchable, and easily-to-navigate interface for users to ask for help Accordingly, we have commissioned
% a Discourse site which is more open than the Facebook group and more user-oriented
% than Astropy Slack. A benefit of the Moore Foundation grant is that Astropy
% developers are able to invoice as independent contractors the time they
% spend helping users on these forums.
% \end{itemize}

@@

- Astropy is established; challenges faced have evolved
-

\begin{itemize}
\item Attracting new contributors when the code has become quite complex, project established (astropy as an "institution", walled-off group). (how to attract? Fast feedback and code reviews, ...)
\item Support for existing contributors: Contributor and maintainer mentoring (...)
\item Infrastructure: Long-term / sustained funding for maintaining infrastructure (grants last 3 years; infrastructure evolves, nature of open source ecosystem)
\item As ecosystem grows, making sure users know about all of the functionality (workshops, tutorials, educational materials, discoverability, accessibility) (Continued engagement with user community)
\end{itemize}

>>>>

Within \astropypkg, development may be roughly split into several
categories: new features, intra- and inter-package interoperability,
improvements to precision and reproducibility, performance and documentation
enhancements, and bug fixes. We discuss some of these in the subsections below.

@@

\astropy now supports reading and writing tables in the American Astronomical
Society Journals' Machine-Readable Table (\href{https://docs.astropy.org/en/stable/io/ascii/write.html#cds-mrt-format}{MRT}) format. This ASCII format has
long been missing and was added to \astropyTable\ I/O in a Google Summer of Code 2021 project.
In addition, \astropyTable\ may also read from and write
to \href{https://asdf-standard.readthedocs.io/en/latest/}{ASDF},
\href{https://docs.astropy.org/en/stable/io/unified.html#table-io-parquet}{Parquet},
and \href{https://wwwastro.msfc.nasa.gov/qdp/}{QDP} formats.

@@

\subsection{Precision \& Reproducibility}
\label{sec:core-features-precision_accuracy_repreducibility}

@@

@@

Community Engagement

\begin{itemize}
\item {\bf User forums:} The Astropy Project has historically maintained a number of avenues for users to seek help or gain access to the developer community. This includes the Astropy mailing list, the Python in Astronomy Facebook group, and the Astropy Slack workspace. The Moore Foundation grant allows this Slack space to
be on a paid plan; additionally NumFOCUS has negotiated a special rate for
open-source projects. While the Slack and Astropy-dev mailing lists are primarily used to discuss the project direction and updates, it was noted that the use of a Facebook community could present a barrier to open source. The Astropy Project identified a need for a public, archived, searchable, and easy-to-navigate interface for users to ask for help. Accordingly, we have commissioned
a Discourse site which is more open than the Facebook group and more user-oriented
than Astropy Slack. A benefit of the Moore Foundation grant is that Astropy
developers are able to invoice as independent contractors the time they
spend helping users on these forums.
\end{itemize}

\begin{itemize}
\item Attracting new contributors when the code has become quite complex,
\item Contributor to maintainer mentoring,
\item Long-term / sustained funding for maintaining infrastructure,
\item ...
\end{itemize}
